Symmetric Circuits for Rank Gates

### 2018-02-05 sections/translating-formulas-to-cicuits.tex
Feb 5

\documentclass[../paper.tex]{subfiles}
\begin{document}
In this section we first introduce a number of useful polynomial-time computable
functions which define transformations on circuits. In the second subsection we
use these these functions to define a transformation of $\FPR$-formulas into
equivalent families of transparent rank-circuits.

@@

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma , \Lambda, L \rangle$ be a circuit and $H
  := \{h_1, \ldots, h_p\}$ be a set of syntactically equivalent gates. We let
  $\op{merge}(C, H) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{h_2, \ldots, h_p\}$,
  \item $\Omega' = \Omega$,
  \item $\Sigma' = \restr{\Sigma}{G'}$,
  \item $\Lambda' = \Lambda$, and
  \item For all $g \in G'$ and $\vec{a} \in Ind(g)$ if $L(g) (\vec{a}) \in H$
    then $L'(g)(\vec{a}) = h_1$ and if $L(g)(\vec{a}) \notin H$ then
    $L'(g)(\vec{a}) = L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

It can be shown that $\op{merge}$ can be computed in time polynomial in the size
of the input circuit. We note that the input and output circuits compute the
same function. Moreover, if the input circuit is symmetric (resp. has unique
labels, is transparent) then the output circuit will also be symmetric (resp.
have unique labels, is transparent).

@@

We can compute $\op{merge-all}$ using the following algorithm. Let $(C, H)$ be a
tuple in the domain of $\op{merge-all}$. We assume a reasonable encoding of the
circuit, and note that any such encoding induces a linear order on the
structure. Let $H_1, \ldots, H_p$ be the syntactic equivalence classes of $H$
linearly ordered such that for all $i, j \in [p]$ if $i \leq j$ the height of
the gates in $H_i$ is less than or equal to the height of the gates in $H_j$.
Let $C'$ be the result of successively executing the algorithm for $\op{merge}$
on each such equivalence class in order.

It is worth noting that for an input tuple $(C, H)$, if follows from
Lemma~\cite{lem:unique-labels-syntactic-equiv} that if $H$ has unique labels
then the syntactic equivalence relation on $H$ can be computed in time
polynomial in the size of $C$. Thus, if the set $H$ has unique labels, it can be
shown that the operation $\op{merge-all}$ can be computed in time polynomial in
the size of the input circuit.

As in the case of $\op{merge}$, the function $\op{merge-all}$ preserves
symmetry, unique labels and the ensures that the input and output circuits
compute the same function. Let be a circuit $C$ a circuit and $H$ a set of gates
in $C$ with unique labels. Then if $g$ is a gate in both $C$ and $C' :=
\op{merge-all}(C, H)$, and all of the children of $g$ in $C$ are in $H$, then no
two distinct children of $g$ in $C'$ are syntactically-equivalent. We note we
have no guarantee that $g$ will have unique labels in $C'$ as $L'(g)$ may not be
injective.

@@

\begin{definition}
  \label{def:make-injective}
  The function $\op{make-injective}$ maps a circuit $C := \langle G, \Omega,
  \Sigma , \Lambda, L \rangle$ and a gate $h \in G$ to the circuit
  $\op{make-injective}(C, h) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, defined as follows. Let $c = \max \{\vert L(g)^{-1}(h)\vert : g \in
  W(h, \cdot)\}$. If $c = 1$ then $C' := C$. If $c > 1$ then $C'$ is defined
  such that
  \begin{itemizens}
  \item $G' = G \uplus \{h_2, \ldots, h_{c-1}\}$,
  \item $\Omega' = \Omega$,
  \item For each $g \in G$ let $\Sigma' (g) = \Sigma(g)$, otherwise let
    $\Sigma'(g)$ be a single-input $\AND$-gate.
  \item $\Lambda'= \Lambda$
  \item For each $g \in G'$ if $g \in G \setminus W(h, \cdot)$ then $L'(g) =
    L(g)$. If $g \in W(h, \cdot)$ let $\{\vec{a}_1, \ldots, \vec{a}_k\} :=
    L(g)^{-1}(h)$ and for each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) \neq h$
    then $L'(g)(\vec{a}) = L(g)(\vec{a})$ and if $\vec{a} = \vec{a_i} \in
    L(g)^{-1}(h)$ then let $L'(g)(\vec{a}) = h_i$.
  \end{itemizens}
\end{definition}

@@

It can be shown that $\op{insert}$ can be be computed in time polynomial in the
size of the input circuits. Moreover, if the input circuits are symmetric then
the output circuit will be symmetric.

@@

\begin{definition}
  The function $\op{replace}$ maps a circuit $C = \langle G, \Omega, \Sigma,
  \Lambda, L\rangle$ and a pair of gates $(g_1, g_2)$ in $C$ to a circuit
  $\op{replace}(C, g_1, g_2) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{g_1\}$,
  \item $\Omega'$ is an function from the domain of $\Omega$ to $G'$ such that
    for each $\vec{a} \in \dom(\Omega)$ if $\Omega(\vec{a}) = g_1$ then
    $\Omega'(\vec{a}) = g_2$ and otherwise $\Omega'(\vec{a}) = \Omega(\vec{a})$.
  \item $\Sigma' = \restr{\Sigma}{G'}$
  \item $\Lambda' = \restr{\Lambda}{G'}$
  \item For each $g \in G'$ and each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) =
    g_1$ then $L'(g)(\vec{a}) = g_2$ and if not then $L'(g)(\vec{a}) =
    L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

@@

\begin{definition}
  The function $\op{tag}$ maps a $(\mathbb{B}, \tau)$-circuit $C := \langle G,
  \Omega, \Sigma, \Lambda, L \rangle$ of order $n$, a relation symbol $R \in
  \tau$, a gate $g \in G$, and a tuple $\vec{a} \in [n]^k$ for some $k$ to a
  circuit $\op{tag}(C, R, g, \vec{a}) := C' := \langle G', \Omega', \Sigma',
  \Lambda', L'\rangle$ of order $n$, such that $G' = G \uplus \{\land_g,
  \land_{g'}, \land_1, \ldots , \land_k\}$, $\Omega' = \Omega$, and $\Lambda' =
  \Lambda$. For each $g \in G$ $\Sigma'(g) = \Sigma(g)$. For each $g \in
  \{\land_0, \land_1, \ldots , \land_k \}$, $\Sigma(g)$ is an $\AND$-gate that
  takes in two inputs. $\Sigma(\land_g)$ is an $\AND$-gate that takes in a
  single input, $g$ is the lone input to $\land_g$, and $\land_0$ has as input
  $\land_g$ and $\land_1$. For each $i \in [k-1]$ the inputs of $\land_i$ are
  $\land_{i+1}$ and the gate $g_i \in G'$ such that $\Sigma(g_i) = R$ and
  $\Lambda^{-1}(g_i) = \vec{a}_i$, where $\vec{a}_i$ is the $k$-tuple
  $(\vec{a}(i), \ldots, \vec{a}(i))$. The inputs of $\land_k$ are the constant
  gate $1$ and the gate $g_k$ such that $\Sigma(g) = R$ and $\Lambda^{-1}(g_k) =
  \vec{a}_k$. For each gate $h \in G \setminus \{g\}$ and $\vec{b} \in \ind(h)$,
  if $L(h)(\vec{b}) = g$ then $L'(h)(\vec{b}) = \land_0$, otherwise
  $L'(h)(\vec{b}) = L(h)(\vec{b})$.
\end{definition}

@@

We now define the function $\op{compose}$ that maps a circuit and a sequence of
appropriate circuits and to a circuit formed by replacing the input gates of the
designated circuit with the output gates of the circuits in the sequence of
circuits.

@@

  Let ${C^0}' := \langle G_0', \Omega_0', \Sigma_0', \Lambda_0', L_0' \rangle :=
  C$, and let $i \in [t]$. Let $D^{i-1}:= \op{insert} (C^i, {C^{i-1}}')$ and
  $G^{i-1}_{R_i}$ be the set of all relational gates in $D^{i-1}$ labelled by
  the relational symbol in $R_i$. Let ${C^{i-1}}' := \langle G_i', \Omega_i',
  \Sigma_i', \Lambda_i', L_i' \rangle$ be the circuit formed from $D^{i-1}$ by
  replacing each gate in $g \in G^{i-1}_{R_i}$ with the gate
  $\Omega_{i}(\Lambda_{i-1}'(g))$ in $D^{i-1}$ (by repeated calls to the
  $op{replace}$ function). Let $C' = {C^{t}}'$.
\end{definition}

It can be shown that $\op{compose}$ can be computed in time polynomial in the
sum of the sizes of the input circuits. Moreover, if the input circuits are all
symmetric then the output circuit will be symmetric. Also, for an input tuple
$(C, \mathcal{C}, \tau_R)$, we have that if $C$ has symmetric gates and
transparent, and all the circuits in $\mathcal{C}$ are transparent, then $C' :=
\op{compose}(C, \mathcal{C}, \tau_R)$ is transparent. If $C$ does not have
symmetric gates then, as the output gates of the circuits in $\mathcal{C}$ are
no longer labelled as output gates in $C'$, it may be that some of these gates
are syntactically-equivalent in $C'$, resulting in $C'$ not being transparent.

In certain contexts we are interested composing gates such that if the input
circuits are transparent (resp. have unique labels) then the output circuits
will be transparent (resp. have unique labels). In this case, while composing
the circuits we could apply the $\op{tag}$ function to the output gates of the
circuits in the input sequence. We now define the function
$\op{compose-and-tag}$ that implements this mapping.

@@

It is easy to see that $\op{compose-and-tag}$ can be computed in time polynomial
in the sun of the sizes of the input circuits. We also have that if the input
circuits are symmetric then the output circuits are symmetric. Furthermore, if
the input sequences of circuits and relations are both singletons, then if the
input circuits are all transparent (resp. have unique labels) then the output
circuit is transparent (resp. have unique labels).

@@

  % We also note that this construction gives us that, if we exclude the cost of
  % the recrive calls, then $n \mapsto T_n(\psi(\vec{y}))$ is computable in time
  % polynomial in $n$. Notice, though, that this polynomial does depend on
  % $\psi'(\vec{y})$, i.e. the form of $\psi (\vec{y})$.

@@

  The above definition can be used to define a recursive algorithm computing the
  function $n \mapsto T_n (\theta_n(\vec{x}))$. We note that, since
  $\width(\theta_n) \leq 4 t$ for all $n \in \nats$, there are only a finite
  number of cases that need to be considered at each recursive step in this
  algorithm. This fact can be used to show that there exists a polynomial bound
  on the running time of the algorithm that depends only on $n$ and the
  structure of $\theta(\vec{x})$. As such, we have that $(T_n (\theta_n))_{n \in
    \nats}$ translates $\theta(\vec{x})$, and the result follows.

  % We first construct an auxiliary formula $\psi'(\vec{y}$ in which all the
  % sub-formulas of $\psi(\vec{y})$ are replaced with a relation symbol. We then
  % construct a circuit $C^{\psi'}$ translating $\psi'(\vec{y})$ at $n$ (either
  % explicitly, as in the rank case, or using the translation defined by
  % Anderson and Dawar~\cite{}). Finally, we compose the circuit $C^{\psi'}$
  % with the recursively defined circuits translating the sub-formulas of
  % $\psi(\vec{y})$ (when composing in the rank case case we add an additional
  % gadget, but the computation cost of this function is similar).

  % The total computation cost of computing $T_n(\psi(\vec{y}))$ is then the
  % cost of constructing $\psi'(\vec{y})$ + the cost of constructing
  % $C^{\psi'}_n$ + the cost of the recursive calls + the cost of composition.
  % We notice that, since there are only a finite number of cases, there are
  % only a finite number of possible formulas $\psi'(\vec{y})$. Let
  % $q_{\psi}(n)$ be a polynomial bounding time required to compute $n \mapsto
  % C^{\psi'}_n$ for the case of $\psi(\vec{y})$. There are only a finite number
  % of cases, so we can define $q_M(n)$ such that $q_M(n) \geq q_\psi(n)$ for
  % all $n \in \nats$ and all cases $\psi(\vec{y})$. Let $q_c(x)$ be a
  % polynomial larger than the polynomials bounding the run time of
  % $\op{compose}$ and $\op{\compose-and-tag}$. Let $p(n)$ be the polynomial
  % bounding witnessing the $P$-uniformity of $(\theta_n)_{n \in \nats}$. We
  % then have that
  % \begin{align*}
  %   \time {T_n(\psi(\vec{y}))} := q_c(\vert C^{\psi'}_n \vert + \vert
  %   %   T_n(\phi_1(\vec{y})) \vert + \vert (\phi_2(\vec{y}))) + \time {T_n
  %   %   (\phi_1(\vec{y}))} + \time {T_n (\phi_2(\vec{y}} + q_M(n)
  % \end{align*}

  % For $i \in [2]$, it can be shown that
  % \begin{align*}
  %   \vert T_n(\phi_i) \vert  =  
  % \end{align*}
  % Let $p(n)$

  % there exists a polynomial $q(n)$ such that $n \maps T_n (\psi(\vec{y}))$ is
  % computable in time

  % . Let $p(n)$ be the polynomial witnessing the $P$-uniformity of
  % $(\theta_n)_{n
  % \in \nats}$. We begin with the bases cases. We notice that in each case

  % We now show that $T_n$ can be computed in time polynomial in $n$, thus
  % giving
  % us that $n \mapsto T_n (\theta_n(\vec{x}))$ can be computed computed in time
  % polynomial in $n$, hence proving the result.

  % It is easy to see that for $\psi (\vec{y})$ a subformula of $\theta_n$, the
  % number of recursive calls needed to define $T_n(\psi(\vec{y}))$ is bounded
  % by
  % the number of sub-formulas in $\theta_n(\vec{x})$, which in turn is bounded
  % by
  % the size of $\theta_n(\vec{x})$, and hence by $p(n)$, where $p$ is the
  % polynomial witnessing the $P$-uniformity of $(\theta_n)_{n \in \nats}$.

  % We note that, up to renaming of variables, there are a constant number of
  % cases to consider (i.e. possible forms of $\psi (\vec{y})$) at each
  % recursive
  % call (this follows from the fact that $\width(\theta_n) \leq 4\width
  % (\theta)$
  % for all $n \in \nats$).

  % A the polynomial is dependent on the form being considered. Since there are
  % only a finite number of such forms, we can sum up these polynomial bounds
  % and
  % arrive at a polynomial bound

  % can appear as sub-formulas of $\theta_n$.

  % It is easy to see that our definition of $T_n$ requires at most recursive
  % steps

  % We have now completed our definition of $T_n$. In each recursive step of our
  % definition we used the translation of Anderson and Dawar~\ref{}

  % We note that $C^{\psi'}_n$ is a transparent symmetric circuit with the
  % property

  % It is easy to see that for all $n \in \nats$, computes the same query as
  % defined by $\theta(\vec{x})$ for structures of size $n$. Moreover, since
  % $C^{\psi'}_n$ and $C^{\phi}_n$ are symmetric, transparent, and the input
  % sequences of relations and circuits are singletons, it follows that
  % $C^{\theta}_n$ is a transparent symmetric rank-circuit. W have thus defined
  % a
  % mapping from $n \mapsto C^{\theta}_n$, that, and that mapping can be
  % computed
  % in time polynomial in $n$.

  % It remains only to note that any atomic formula of $\FPR[\tau]$ is trivially
  % circuit-translatable. The result thus follows.

  % $p$ is a prime, $r$ is a natural number, and $\theta(\vec{x}) := \rank^{\leq
  % r}_p \vec{z} \vec{y}. \phi(\vec{z}, \vec{y}, \vec{x})$. We define an
  % algorithm that takes as input a natural number $n$ outputs the required
  % circuit $C^{\theta}_n$.

  % The approach used for resolving the conjunctive case can be thought of as
  % having three steps. First, replacing each sub-formula in $\psi$ with a
  % relation symbol, resulting in a formula $\psi'$ of $\FPC[\tau]$. Second,
  % using
  % the translation of Anderson and Dawar to define the corresponding
  % transparent
  % symmetric circuit $C^{\psi'}_n$. Third, recursively computing the circuits
  % for
  % the sub-formulas, and then composing the circuits in order to construct
  % $T(\psi(\vec{y}))$.

  % Notice that if the sub-formulas are circuit-translatable then

  % (and the results of Anderson and Dawar) on $\theta'$ and the sub-formulas,
  % producing appropriate families for $\theta'$ and each sub-formula of
  % $\theta$.
  % Third, for each $n \in \nats$ composing the circuit $C^{\theta'}_n$ with the
  % circuits for each sub-formula, producing the the circuit $C^{\theta}_n$.
  % Since
  % this process can be executed in time polynomial in $n$, we conclude that
  % $\theta$ is circuit-translatable.

  % This same approach can be used to handle many other cases. For
  % circuit-translatable $\phi, \phi_1, \phi_2 \in \FPR[\tau]$, and $k \in
  % \nats$,
  % similar constructions can be used to prove that if $\theta(\vec{x}) \in
  % \FPR[\tau]$ is of the form $\phi_1(\vec{x}) \lor \phi_2(\vec{x})$, $\neg
  % \phi(\vec{x})$, $\forall y. \phi (\vec{x}, y)$, $\exists y. \phi (\vec{x},
  % y)$
  % or $C^k y. \phi(\vec{x}, y)$ then $\theta(\vec{x})$ is circuit-translatable.

  % It follows that the number of sub-formulas of $\theta_n$

  % It follows that in order to prove that $\theta(\vec{x})$ is
  % circuit-translatable it is sufficient to prove that there exists a function
  % that maps $\theta_n$ (for any $n \in \nats$) to $C^{\theta}_n$, a
  % transparent
  % symmetric rank-circuit that computes the same query as defined by
  % $\theta_n(\vec{x})$ (and so $\theta(\vec{x})$) for structures of size $n$,
  % and
  % is computable in time polynomial in in both $n$ and the number of
  % sub-formulas
  % in $\theta_n$. This is sufficient as the number of sub-formulas in
  % $\theta_n$
  % is bounded by some polynomial in $n$, and so composing functions, we can
  % compute the map that takes $n$ to $\theta_n$ to $C^{\theta}_n$ in time
  % polynomial in $n$.

  % We define an algorithm for constructing $C^{\theta}_n$ given
  % $\theta_n(\vec{x})$ by recursion on the structure of the input formula. The
  % first note that if $\theta_n(\vec{x})$ an atomic formula then there is a
  % trivial construction of $C^{\theta}_n$.

  % We consider the conjunctive case. Suppose $\phi_1, \phi_2 \in \FPR[\tau]$
  % are
  % circuit-translatable and $\theta (\vec{x}) = \phi_1(\vec{x}) \land
  % \phi_2(\vec{x})$. Let $R_{\phi_1}$ and $R_{\phi_2}$ be two $\vert \vec{x}
  % \vert$-ary relation symbols not in $\tau$. Let $\theta' (\vec{x}) :=
  % R_{\phi_1}(\vec{x}) \land R_{\phi_2}(\vec{x})$. Let $n \in \nats$. Since
  % $\theta'(\vec{x})$ is a formula of $\FPC[\tau]$, we have from Anderson and
  % Dawar~\cite{AndersonD17} that $\theta'(\vec{x})$ is circuit-translatable. We
  % let $C^{\theta}_n := \op{compose}(C^{\theta'}, (C^{\phi_1}, C^{\phi_2}),
  % (R_{\phi_1}, R_{\phi_2}))$.

  % Since all internal gates in $C^{\theta'}$ are symmetric, and both
  % $C^{\phi_1}$
  % and $C^{\phi_2}$ are transparent, it follows that $C^{\theta}_n$ is
  % transparent. Moreover, it is easy to see that the construction of
  % $C^{\theta}_n$ can be carried implemented so as to run in time polynomial in
  % $n$, and that $(C^{\theta}_n)_{n \in \nats}$ computes the same query defined
  % by $\theta(\vec{x})$. It follows that $\theta(\vec{x})$ is
  % circuit-translatable.

  % The approach used for resolving the conjunctive case can be thought of as
  % having three steps. First, replacing each sub-formula in $\theta$ with a
  % relation symbol, resulting in a formula $\theta'$ of $\FPC$. Second, using
  % the
  % algorithms from the inductive hypothesis' (and the results of Anderson and
  % Dawar) on $\theta'$ and the sub-formulas, producing appropriate families for
  % $\theta'$ and each sub-formula of $\theta$. Third, for each $n \in \nats$
  % composing the circuit $C^{\theta'}_n$ with the circuits for each
  % sub-formula,
  % producing the the circuit $C^{\theta}_n$. Since this process can be executed
  % in time polynomial in $n$, we conclude that $\theta$ is
  % circuit-translatable.

  % This same approach can be used to handle many other cases. For
  % circuit-translatable $\phi, \phi_1, \phi_2 \in \FPR[\tau]$, and $k \in
  % \nats$,
  % similar constructions can be used to prove that if $\theta(\vec{x}) \in
  % \FPR[\tau]$ is of the form $\phi_1(\vec{x}) \lor \phi_2(\vec{x})$, $\neg
  % \phi(\vec{x})$, $\forall y. \phi (\vec{x}, y)$, $\exists y. \phi (\vec{x},
  % y)$
  % or $C^k y. \phi(\vec{x}, y)$ then $\theta(\vec{x})$ is circuit-translatable.

  % Suppose $\phi(\vec{z}, \vec{y}, \vec{x}) \in \FPR[\tau]$ is
  % circuit-translatable, $p$ is a prime, $r$ is a natural number, and
  % $\theta(\vec{x}) := \rank^{\leq r}_p \vec{z} \vec{y}. \phi(\vec{z}, \vec{y},
  % \vec{x})$. We define an algorithm that takes as input a natural number $n$
  % outputs the required circuit $C^{\theta}_n$.

  % From the inductive hypothesis we may efficiently compute $C^{\phi}_n =
  % \langle
  % G^{\phi}, \Omega^{\phi}, \Sigma^{\phi}, \Lambda^{\phi}, L^{\phi} \rangle$.
  % Let
  % $k:= \vert \vec{x} \vert$, $k_1 : = \vert \vec{z} \vert$ and $k_2 := \vert
  % \vec{y} \vert$. Let $R_\phi$ be a $(k_1 + k_2 + k)$-ary relation symbol not
  % in
  % $\tau$.

  % Let $\theta'(\vec{x}) = \rank^{\leq r}_r := \vec{z}, \vec{y}. R_{\phi}
  % (\vec{z}, \vec{y}, \vec{x})$. We define the $(\mathbb{B}_{\rank}, \{
  % R_{\phi}
  % \})$-circuit $C^{\theta'}_n:= \langle G^{\theta'}, \Omega^{\theta'},
  % \Sigma^{\theta'}, \Lambda^{\theta'}, L^{\theta'}\rangle$ as follows. Let
  % $G_{\text{out}} := \{g_{\vec{c}} : \vec{c} \in [n]^k\}$ and $G_{\text{in}} :
  % =
  % \{ g_{\vec{a}, \vec{b}, \vec{c}} : \vec{a} \in [n]^{k_1}, \vec{b} \in
  % [n]^{k_2}, \vec{c} \in [n]^{k}\}$.

  % \begin{itemize}
  %   \setlength\itemsep{0mm}
  % \item Let $G^{\theta'} = G_{\text{out}} \uplus G_{\text{in}}$,
  % \item $\Omega^{\theta'}$ be an injection from $[n]^k$ to $G^{\theta'}$
  %   defined
  %   by $\Omega^{\theta'}(\vec{c}) = g_{\vec{c}}$ for all $\vec{c} \in [n]^k$,
  % \item $\Sigma^{\theta'} (g) = \rank^r_p$ if $g \in G_{\text{out}}$ and
  %   $\Sigma^{\theta'}(g) = R_\phi$ if $g \in G_{\text{in}}$,
  % \item For each $(\vec{a}, \vec{b}, \vec{c}) \in [n]^{k_1 +k_2 + k}$ let
  %   $\Lambda^{\theta'} (\vec{a}, \vec{b}, \vec{c}) := g_{\vec{a}, \vec{b},
  %   \vec{c}}$, and
  % \item For each gate $g_{\vec{c}} \in G_{\text{out}}$ let $L(g_{\vec{c}}):
  %   [n]^{k_1} \times [n]^{k_2} \rightarrow G^{\theta'}$ be defined by
  %   $L(g_{\vec{c}})(\vec{a}, \vec{b}) := g_{\vec{a}, \vec{b}, \vec{c}}$ for
  %   each
  %   $\vec{a} \in [n]^{k_1}$ and $\vec{b} \in [n]^{k_2}$.
  % \end{itemize}

  % % It is easy to see that both $C^{\theta'}_n$ (and $C^{\phi}_n$) are
  % % symmetric
  % % rank-circuits with unique labels.

  % % It is easy to see that every gate in $$

  % % be the result of calling $\op{tag}$ on

>>>>

\documentclass[../paper.tex]{subfiles}
\begin{document}
In this section we first introduce a number of useful polynomial-time computable
functions which define transformations on circuits. In the second subsection we
use these functions to define a transformation that takes $\FPR$-formulas to
equivalent families of transparent symmetric rank-circuits.

@@

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma , \Lambda, L \rangle$ be a circuit and $H
  := \{h_1, \ldots, h_p\}$ be a set of syntactically equivalent gates. We let
  $\op{merge}(C, H) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{h_2, \ldots, h_p\}$,
  \item $\Omega' = \Omega$,
  \item $\Sigma' = \restr{\Sigma}{G'}$,
  \item $\Lambda' = \Lambda$, and
  \item For all $g \in G'$ and $x \in Ind(g)$ if $L(g) (x) \in H$ then $L'(g)(x)
    = h_1$ and if $L(g)(x) \notin H$ then $L'(g)(x) = L(g)(x)$.
  \end{itemize}
\end{definition}

It can be shown that $\op{merge}$ can be computed in time polynomial in the size
of the input circuit. We note that the input and output circuits compute the
same function. Moreover, if the input circuit is symmetric (resp. has unique
labels, is transparent) then the output circuit will also be symmetric (resp.
have unique labels, transparent).

@@

We can compute $\op{merge-all}$ using the following algorithm. Let $(C, H)$ be a
tuple in the domain of $\op{merge-all}$. We assume a reasonable encoding of the
circuit, and note that any such encoding induces a linear order on the
structure. Let $H_1, \ldots, H_p$ be the syntactic-equivalence classes of $H$
linearly ordered such that for all $i, j \in [p]$ if $i \leq j$ the height of
the gates in $H_i$ is less than or equal to the height of the gates in $H_j$.
Let $C'$ be the result of successively executing the algorithm for $\op{merge}$
on each such equivalence class in order.

It is worth noting that for an input tuple $(C, H)$, if follows from
Lemma~\cite{lem:unique-labels-syntactic-equiv} that if $H$ has unique labels
then the syntactic-equivalence relation on $H$ can be computed in time
polynomial in the size of $C$. Thus, if each gate in $H$ has unique labels, it
can be shown that $\op{merge-all}$ can be computed in time polynomial in the
size of the input circuit.

As in the case of $\op{merge}$, the function $\op{merge-all}$ preserves
symmetry, unique labels and the ensures that the input and output circuits
compute the same function. Let be a circuit $C$ a circuit and $H$ a set of gates
in $C$ with unique labels. Then if $g$ is a gate in both $C$ and $C' :=
\op{merge-all}(C, H)$, and all of the children of $g$ in $C$ are in $H$, then no
two distinct children of $g$ in $C'$ are syntactically-equivalent. However, we
have no guarantee that $g$ will have unique labels in $C'$ as we have no
guarantee that $L'(g)$ is injective.

@@

\begin{definition}
  \label{def:make-injective}
  The function $\op{make-injective}$ maps a circuit $C := \langle G, \Omega,
  \Sigma , \Lambda, L \rangle$ and a gate $h \in G$ to the circuit
  $\op{make-injective}(C, h) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, defined as follows. Let $c = \max \{\vert L(g)^{-1}(h)\vert : g \in
  W(h, \cdot)\}$. If $c = 1$ then $C' := C$. If $c > 1$ then $C'$ is defined
  such that
  \begin{itemizens}
  \item $G' = G \uplus \{h_2, \ldots, h_{c-1}\}$,
  \item $\Omega' = \Omega$,
  \item For each $g \in G$ let $\Sigma' (g) = \Sigma(g)$, otherwise let
    $\Sigma'(g)$ be a single-input $\AND$-gate.
  \item $\Lambda'= \Lambda$
  \item For each $g \in G'$ if $g \in G \setminus W(h, \cdot)$ then $L'(g) =
    L(g)$. If $g \in W(h, \cdot)$ let $\{x_1, \ldots, x_k\} := L(g)^{-1}(h)$ and
    for each $x \in \ind(g)$ if $L(g)(x) \neq h$ then $L'(g)(x) = L(g)(x)$ else
    if $x = x_i \in L(g)^{-1}(h)$ then let $L'(g)(x) = h_i$.
  \end{itemizens}
\end{definition}

@@

It can be shown that $\op{insert}$ can be be computed in time polynomial in the
size of the input circuits. Moreover, if the input circuits are symmetric then
the output circuit is symmetric.

@@

\begin{definition}
  The function $\op{replace}$ maps a circuit $C = \langle G, \Omega, \Sigma,
  \Lambda, L\rangle$ and a pair of gates $(g_1, g_2)$ in $C$ to a circuit
  $\op{replace}(C, g_1, g_2) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{g_1\}$,
  \item $\Omega'$ is an function from the domain of $\Omega$ to $G'$ such that
    for each $\vec{a} \in \dom(\Omega)$ if $\Omega(\vec{a}) = g_1$ then
    $\Omega'(\vec{a}) = g_2$ and otherwise $\Omega'(\vec{a}) = \Omega(\vec{a})$.
  \item $\Sigma' = \restr{\Sigma}{G'}$
  \item $\Lambda' = \restr{\Lambda}{G'}$
  \item For each $g \in G'$ and each $x \in \ind(g)$ if $L(g)(x) = g_1$ then
    $L'(g)(x) = g_2$ and if not then $L'(g)(x) = L(g)(x)$.
  \end{itemize}
\end{definition}

@@

\begin{definition}
  The function $\op{tag}$ maps a $(\mathbb{B}, \tau)$-circuit $C := \langle G,
  \Omega, \Sigma, \Lambda, L \rangle$ of order $n$, a relation symbol $R \in
  \tau$, a gate $g \in G$, and a tuple $\vec{a} \in [n]^k$ for some $k$ to a
  circuit $\op{tag}(C, R, g, \vec{a}) := C' := \langle G', \Omega', \Sigma',
  \Lambda', L'\rangle$ of order $n$, such that $G' = G \uplus \{\land_g,
  \land_{g'}, \land_1, \ldots , \land_k\}$, $\Omega' = \Omega$, and $\Lambda' =
  \Lambda$. For each $g \in G$ $\Sigma'(g) = \Sigma(g)$. For each $g \in
  \{\land_0, \land_1, \ldots , \land_k \}$, $\Sigma(g)$ is an $\AND$-gate that
  takes in two inputs. $\Sigma(\land_g)$ is an $\AND$-gate that takes in a
  single input, $g$ is the lone input to $\land_g$, and $\land_0$ has as input
  $\land_g$ and $\land_1$. For each $i \in [k-1]$ the inputs of $\land_i$ are
  $\land_{i+1}$ and the gate $g_i \in G'$ such that $\Sigma(g_i) = R$ and
  $\Lambda^{-1}(g_i) = \vec{a}_i$, where $\vec{a}_i$ is the $k$-tuple
  $(\vec{a}(i), \ldots, \vec{a}(i))$. The inputs of $\land_k$ are the constant
  gate $1$ and the gate $g_k$ such that $\Sigma(g) = R$ and $\Lambda^{-1}(g_k) =
  \vec{a}_k$. For each gate $h \in G \setminus \{g\}$ and $x \in \ind(h)$, if
  $L(h)(x) = g$ then $L'(h)(x) = \land_0$, otherwise $L'(h)(x) = L(h)(x)$.
\end{definition}

@@

We now define the function $\op{compose}$ that maps a circuit and a sequence of
circuits and to a circuit formed by replacing the input gates of the designated
circuit with the output gates of the circuits in the sequence.

@@

  Let ${C^0}' := \langle G_0', \Omega_0', \Sigma_0', \Lambda_0', L_0' \rangle :=
  C$, and let $i \in [t]$. Let $D^{i-1}:= \op{insert} (C^i, {C^{i-1}}')$ and
  $G^{i-1}_{R_i}$ be the set of all relational gates in $D^{i-1}$ labelled by
  the relational symbol in $R_i$. Let ${C^{i-1}}' := \langle G_i', \Omega_i',
  \Sigma_i', \Lambda_i', L_i' \rangle$ be the circuit formed from $D^{i-1}$ by
  replacing each gate in $g \in G^{i-1}_{R_i}$ with the gate
  $\Omega_{i}(\Lambda_{i-1}'(g))$ in $D^{i-1}$ (by repeated calls to the
  $\op{replace}$ function). Let $C' = {C^{t}}'$.
\end{definition}

It can be shown that $\op{compose}$ can be computed in time polynomial in the
sum of the sizes of the input circuits. We also note that the size of $C'$ is
less than or equal to the sum of the sizes of the input circuits. Moreover, if
the input circuits are all symmetric then the output circuit will be symmetric.
Also, for an input tuple $(C, \mathcal{C}, \tau_R)$, we have that if $C$ has
symmetric gates and transparent, and all the circuits in $\mathcal{C}$ are
transparent, then $C' := \op{compose}(C, \mathcal{C}, \tau_R)$ is transparent.
If $C$ contains a non-symmetric gates then, as the output gates of the circuits
in $\mathcal{C}$ are no longer labelled as output gates in $C'$, it may be that
some of these gates are syntactically-equivalent in $C'$, resulting in $C'$ not
being transparent.

In certain contexts we are interested composing gates such that if the input
circuits are transparent (resp. have unique labels) then the output circuits
will be transparent (resp. have unique labels). In this case, while composing
the circuits we apply the $\op{tag}$ function to the output gates of the
circuits in the input sequence of circuits. We now define the function
$\op{compose-and-tag}$ that implements this operation.

@@

It is easy to see that $\op{compose-and-tag}$ can be computed in time polynomial
in the sun of the sizes of the input circuits. We also have that if the input
circuits are symmetric then the output circuits are symmetric. Furthermore, if
the input sequences of circuits and relations are both singletons, then if the
input circuits are all transparent (resp. have unique labels) then the output
circuit is transparent (resp. have unique labels). We also note that the output
circuit

@@

@@

  The above definition can be used to define a recursive algorithm computing the
  function $n \mapsto T_n (\theta_n(\vec{x}))$. We note that, since
  $\width(\theta_n) \leq 4 t$ for all $n \in \nats$, there are only a finite
  number of cases that need to be considered at each recursive step in this
  algorithm. This fact can be used to show that there exists a polynomial bound
  on the running time of the algorithm that depends only on $n$ and the
  structure of $\theta(\vec{x})$. As such, we have that $(T_n (\theta_n))_{n \in
    \nats}$ translates $\theta(\vec{x})$, and the result follows.
\end{proof}

\bibliography{../references.bib}
% \ifcsdef{mainfile}{}{\bibliography{../references.bib}}
\end{document}
